---
title: "32061412_Assignment2"
author: "HongYi"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  html_document:
    code_folding: show
    highlight: haddock
    theme: flatly
    toc: yes
    toc_float: yes
  pdf_document:
    toc: yes
editor_options: 
  markdown: 
    wrap: sentence
---

## Libraries loading

```{r}
library(tidyverse)
library(lubridate)
```

## Data exploration

```{r}
# read the news data
news <- read_csv("ireland_news.csv")
```

```{r}
head(news)
```

```{r}
glimpse(news)
```

## Question 1

What are the earliest and latest articles from Irish Independent, irrespective of headline category?
Please also sort the data according to the column publish_date in an ascending manner and display the last 5 records of the data.

### Answer

-   only publish_date and headline category are the selected columns because we already know the data is from Irish Independent and headline category is irrelevant

-   before sorting, filter out the NA publish_date column

-   arrange the data based on the publish_date in date type

```{r}
# get Irish Independent's articles only
news_df <- news %>%
  data.frame() %>%
  filter(news_provider == "Irish Independent") %>%
  select(publish_date, headline_text)

# filter out NA date, change the data type of date, and sort asc
news_df <- news_df %>%
  filter(!is.na(publish_date)) %>%
  mutate(publish_date = dmy(publish_date)) %>%
  arrange(publish_date)
  
# view data in ascending order
news_df

# display last 5 records(5 latest data)
tail(news_df, 5)
```

## Question 2

How many unique headline_category values are there in the data file?
Please consider variations (e.g.: capitalisation, potential inconsistencies) of the headline_category values, when counting them.

How many news category articles contain either the keyword, "Ireland", "Irish", "US", or "USA" along with year digits from 2000 to 2024 in headline_text?
For example, you need to search and count articles containing both “Ireland” and the year digits, or containing both “Irish” and the year digits, and so on.

## Question 3

Please display the top 10 headline categories with the largest number of articles published on Monday throughout the years.
Then, draw a chart showing the total number of articles for the top 10 headline categories (as identified previously) for each year.
What can you observe?
Please discuss the chart and your findings.

## Question 4

Compute the total number of articles for each headline category and news provider.
Then, use a single R function/command to display the statistical information, i.e., Min, Max, and Mean, of the total number of articles (as computed previously) for each news provider.
Note: You can use multiple functions/commands to get the desired pre-processed data table, but when you compute and display the statistical information, you need to use a single R function/command.

## Question 5

Please compute the total number of articles for each headline category, news provider, and the day of the week.
Then, compute the average number of articles for each news provider and the day of the week, based on the total number of articles computed previously.
After that, please display the day of the week with the highest average number of articles for each provider.
The output data should be structured in the following format.

## Question 6

## Question 7

## Question 8
